Machine Learning Autoencoder Applied to Communication Channels

### 2019-06-21 Rapport/Article/Article.tex
Seance 06/22/2019

\author{
\IEEEauthorblockN{Eduardo Dadalto Camara Gomes\IEEEauthorrefmark{1},
Meryem Benammar\IEEEauthorrefmark{2}}

\IEEEauthorblockA{\IEEEauthorrefmark{1}Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO), Université de Toulouse, 31055 Toulouse, FRANCE\\
Email: eduardo.dadalto-camara-gomes@student.isae-supaero.fr
}
\IEEEauthorblockA{\IEEEauthorrefmark{2}Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO), Université de Toulouse, 31055 Toulouse, FRANCE\\
Email: meryem.benammar@isae-supaero.fr
}
}

\IEEEspecialpapernotice{(Final report)}

@@

Communication channel error correction is key to enable digital critical communication systems to work efficiently. The maximum a posteriori (MAP) decoder is proven to be the mathematically optimal solution for the problem. However, its implementation cripples the system applicability. This is because the MAP rule introduces a large delay for long code words. In this context, a deep neural network (DNN) is proposed to optimize the channel with an end-to-end autoencoder. The trained DNN autoencoder has a \textit{one-shot} capability which outperforms a MAP decoder for a given encoder in terms of delay. ONE PHRASE FOR RESULTS

@@

We observed during training, while experimenting with different architecture, that the array autoencoders would have a BER performance which looked like the no decoding curve, but translated in the y axis. This phenomenon can be observed in figure \ref{fig:parrayanalysis} where we tested several $p_t$ options to find the best one for implementing a more refined training. It is remarkable how the curves tend to replicate the same type of behavior. I believe it shows how the local minimums of the solution space of a NN are indeed close to each other, and the weight values may not be indeed that different.

@@

\ref{fig:1Hautoencoder}

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
    \caption{One-hot autoencdoer BER performance. Trained without a noise channel.}\label{fig:1Hautoencoder}
\end{figure}

@@

In order to analyze the delay in the communication for each technique studied, a monte carlo simulation was conducted. There were selected 20 different BSC channels. For each channel, $100000$ messages were codified, transmitted and decodified for each decoder, autoencoder and MAP. This process was timed and repeated several times. In the end, an average time and standard deviation of the results were recorded. A normalization for the average MAP decoding time was done. The results are available in the table \ref{tab:timeanalysis}.

@@

\begin{table}
\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the average MAP algorithm decoding time. } 
\begin{center}

@@

@@

Talk about feasibility of decoders, improved performance of the channel delay with DNN decoders, challenges in implementing this architecture in real life, challenges in training the NNs, Future work, what else could be done with NN in error correction, Advantages and disadvantages in comparison with linear decoding.

\section*{Acknowledgment}
The authors would like to thank...

\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18, DBLP:journals/corr/KeskarMNST16, DBLP:journals/corr/IoffeS15,2018arXiv181103378N}

>>>>

\author{
\IEEEauthorblockN{Eduardo Dadalto Camara Gomes\IEEEauthorrefmark{1}}

%%Meryem Benammar\IEEEauthorrefmark{2}}

\IEEEauthorblockA{\IEEEauthorrefmark{1}Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO), Université de Toulouse, 31055 Toulouse, FRANCE\\
Email: eduardo.dadalto-camara-gomes@student.isae-supaero.fr
}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO), Université de Toulouse, 31055 Toulouse, FRANCE\\
%Email: meryem.benammar@isae-supaero.fr
%}
}

%\IEEEspecialpapernotice{(Final report)}

@@

Communication channel error correction is key to enable digital critical communication systems to work efficiently. The maximum a posteriori (MAP) decoder is proven to be the mathematically optimal solution for the problem. However, its implementation cripples the system applicability. This is because the MAP rule introduces a large delay for long code words. In this context, a deep neural network (DNN) is proposed to optimize the channel with an end-to-end autoencoder. The trained DNN autoencoder has a \textit{one-shot} capability which outperforms a MAP decoder for a given encoder in terms of delay. The decoders proved to be around 25\% faster than the MAP and they achieved the same BER performance. There are still opportunities for enhancement for the autoencoders.

@@

We observed during training, while experimenting with different architecture, that the array autoencoders would have a BER performance which looked like the no decoding curve, but translated in the $y$ axis. This phenomenon can be observed in figure \ref{fig:parrayanalysis} where we tested several $p_t$ options to find the best one for implementing a more refined training. It is remarkable how the curves tend to replicate the same type of behavior. I believe it shows how the local minimums of the solution space of a NN are indeed close to each other, and the weight values may not be indeed that different.

@@

The one-hot autoencoder could not outperform the MAP traced for any value of $p$. As observed in Fig. \ref{fig:1Hautoencoder}, even though it learned a similar encoding, the resulting BER estimation is not optimal. Equally to the array autoencoder, we tested a MAP rule for the learned encoder and found that the DNN decoder portion performs just like the MAP. Thus, once again the problem is in the learned encoding. We tried 80 different configurations, though we could not find one which could outperform the reference model. We can also highlight the trace of the BER curve, approximating the reference model one. Other one-hot autoencoders showed similar curves but translated up in the $y$ axis.

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.5\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
    \caption{One-hot autoencdoer BER performance. Trained without a noisy channel. The DNN was trained with $2^{16}$ training epochs.}\label{fig:1Hautoencoder}
\end{figure}

@@

In order to analyze the delay in the communication for each technique studied, a Monte Carlo simulation was conducted. There were selected 20 different BSC channels. For each channel, $100000$ messages were codified, transmitted and decodified for each decoder, autoencoder and MAP. This process was timed and repeated several times. In the end, an average time and standard deviation of the results were recorded. A normalization for the average MAP decoding time was done. The results are available in the table \ref{tab:timeanalysis}.

@@

\begin{table}
\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the average MAP algorithm decoding time. The simulation was based on CPU calculations. } 
\begin{center}

@@

paragraph that talks about CPU, not GPU.

@@

The feasibility of a machine learning based channel decoder and the end-to-end application of a deep neural network autoencoder were demonstrated for a communication system with a binary symmetric channel. Their performances were assessed and analyzed in terms of bit error rate performance and communication delay. A few hyper-parametric analysis were conducted to find the best solution for each technique.

The designed decoders improved around 25\% the channel delay when compared to the maximum a posteriori rule for a $(n,k)$ code with $\textbf{G}$ aforementioned as linear encoding matrix. Their small number of parameters aligned with a GPU equipped decoding computer is promissory to be a substitute of the MAP rule and achieve same bit error correction rate. In addition, their training and architecture are straightforward and can be trained with a commercial computer. Once trained, they work as a non-linear decoder to a specific channel encoding.

We also proved that autoencoders can learn a non linear encoding function and its decoding function for a binary symmetric channel. Thus there is no need for a generator matrix for error correction. There are still some challenges in training to solve, such as ameliorate the architecture so it can outperform the reference model in every channel condition. In addition, because of the number of parameters, their calculation in CPU is slow and they showed a greater delay when compared to the MAP rule. Though, they must take the most advantage of GPU based calculations and the delay time could be improved using this technology.

A more rigorous hyper-parametric analysis could find a combination of parameters which yields better BER performance for the autoencoders and is strongly suggested for future work. In addition, in the future, a real implementation of the DNN decoders could be tested to confirm their capabilities as it is not the main focus of this work. We suggest using more advanced NN topology such as a recurrent neural network (RNN) or a generative adversarial network (GAN). Their implementation and training are more complex, however they have recently demonstrated outstanding results in other fields.

%Talk about feasibility of decoders, improved performance of the channel delay with DNN decoders, challenges in implementing this architecture in real life, challenges in training the NNs, Future work, what else could be done with NN in error correction, Advantages and disadvantages in comparison with linear decoding.

\section*{Acknowledgment}

I would like to thank the guidance and advice from my supervisor Meryem Benammar, Ph.D, my colleague Rémy Zawislak who worked in the same theme and collaborated indirectly to the research and Marjorie Grzeskowiak Lucas, Ph.D, who through the Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO) enabled the confection of the project.

%\cite{Shannon:2001:MTC:584091.584093, DBLP:journals/corr/CalabreseWGPS16, DBLP:journals/corr/OSheaH17, 2016arXiv160806409O, 2017arXiv171008379G, Worm00turbo-decodingwithout, Viterbi, journals/ett/RobertsonHV97, HagenauerJ, JordanMA,Ibnkahla, nielsenneural, murphy2013machine, DBLP:journals/corr/AbadiABBCCCDDDG16, chollet2015keras, doi:10.1162/neco.2006.18.7.1527, DBLP:conf/acssc/BenammarP18, DBLP:journals/corr/KeskarMNST16, DBLP:journals/corr/IoffeS15,2018arXiv181103378N}

### 2019-06-21 Rapport/Presentation/Presentation/Presentation.tex
Seance 06/22/2019

\pgfdeclareimage[height=0.5cm]{university-logo}{images/ISAE}
\logo{\pgfuseimage{university-logo}}

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.8\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
    \caption{One-hot autoencdoer BER performance. Trained without a noise channel.}\label{fig:1hautoencoder}
\end{figure}

@@

\subsection{Time Analysis}
\begin{frame}{Delay Time Analysis}{Comparison of decoding time for each method.}
  \begin{table}
\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the average MAP algorithm decoding time. } 
\begin{center}

@@

\begin{frame}{Acknowledgment}
  \begin{itemize}
  \item {
    My first point.
  }
  \item {
    My second point.
  }
  \end{itemize}
\end{frame}

% All of the following is optional and typically not needed. 
\appendix
\section<presentation>*{Bibliography}

>>>>

%\pgfdeclareimage[height=0.5cm]{university-logo}{images/ISAE}
%\logo{\pgfuseimage{university-logo}}

@@

\begin{figure}[!ht]
  \centering
    \includegraphics[width=0.75\textwidth]{images/MAP_Autoencoder1H_Mep_131072_ptrain_0_192_128_96_64_32-128}
    \caption{One-hot autoencdoer BER performance. Trained without a noisy channel.}\label{fig:1hautoencoder}
\end{figure}

@@

\subsection{Time Analysis}
\begin{frame}{Delay Time Analysis}{Comparison of encoding, transmission and decoding time for each method.}
  \begin{table}
\caption{Decoding time comparison between the MAP algorithm and the DNN decoders and autoencoders. The data is normalized to the average MAP algorithm decoding time. } 
\begin{center}

@@

\begin{frame}{Acknowledgment}
  I would like to thank ...
  \begin{itemize}
   \item{The guidance and advice from my supervisor Meryem Benammar, Ph.D;}
  \item{ My colleague Rémy Zawislak who worked in the same theme and collaborated indirectly to the research;}
   \item{Marjorie Grzeskowiak Lucas, Ph.D, who through the Institut Supérieur de l'Aéronautique et de l'Espace (ISAE-SUPAERO) enabled the confection of the project; and }
   \item{The supervisors, the LACS member and the audience present today.}
  \end{itemize}

\appendix
\section<presentation>*{Bibliography}
